
```{r setup-clipol, results=FALSE, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 7,
                      fig.height = 5,
                      fig.align = 'left')

library(ggplot2)
library(tidyverse)
library(readxl)
library(dplyr)

source('gradu_read_data.R', encoding = 'UTF-8')
```


# Climate policy in Finland {#clipolfi}

In this chapter, I will briefly introduce Finland's climate policy before implementing the European Union emission trading system and briefly introduce the different phases of the European Union emission trading system. The effects of these regulatory updates are discussed in the following chapters. Still, a short synopsis of the regulatory evolution is necessary for the reader to fully appreciate the economic consequences of the policy shocks produced by these institutions. 

## Before the EU Emission trading system {#preeuclipol}

Finland was the first country in the world to implement a tax on the sources of carbon emissions \citep{hallitus1989, bavbek2016}. The tax encompassed various emission sources, as it targeted transport fuels and the fuels used for energy production \citep{lin2011,ekins1999}. The carbon tax was stringent as it was indifferent to the final user and buyer of the energy-producing fuels, as the tax policy did not differentiate between households and industrial users \citep{ekins1999}. The only fuel source exempted from carbon tax was wood \citep{elbaum2021}. The tax was initially modest, with a relatively low valuation of 1.12 € per CO2 equivalent tonne, and the tax was progressively increased to more substantial levels \citep{bavbek2016}. Sweden followed the example laid out by Finland and enacted a carbon tax in 1991 \citep{andersson2019}. In Chapter \@ref(effects), I will elaborate on the effectiveness of the Finnish and Swedish carbon taxes. I will further discuss the literature and the evidence of the effectiveness of different policy choices.

The carbon tax was not implemented as a revenue-raising measure, as it was from the beginning argued from environmental grounds as it could price the externalities caused by the burning of fossil fuels. The government proposal was enacted as law in 1989 and entered into force at the beginning of 1990. \citep{hallitus1989}. 

 The carbon tax policies of the 1990s Nordics might provide an explanation for my counterintuitive findings about the reaction of the Finnish economy to the carbon policy shocks of the European Union, which are discussed in Chapter \@ref(macroeff). The Finnish economy had more time and a solid monetary incentive to use fossil fuel resources more efficiently and to invest in green infrastructure prior to the enaction of the European Union ETS. When compared to other European economies, which did not have such incentives. Even after the enaction of the EU ETS, the tax continues to affect sectors not subject to the trading system \citep{bavbek2016}. 


## EU emission trading system

A new chapter in the Finnish climate policy began in 2005 as the European Union emission trading system (ETS) was established. The ETS is based on a cap-and-trade scheme that restricts the total emissions for the affected sectors (these will vary in the different phases) and lets the market participants trade with each other \citep{garcia2021}. Furthermore,  a variety of allocation schemes have been implemented in the different phases of the EU ETS as the European Commission has updated the allocation strategy \citep{verde2019free}. The ETS has been operating in four phases: Phase I ran from 2005 to 2007 and served as an experimentation period, Phase II was operational from 2008 to 2012, Phase III from 2013 to 2020, and Phase IV will run from 2021 to 2030 \citep{ellerman2020,joltreau2019,verde2019free}.

Phase I of EU ETS is widely considered the experimentation period where the institutions of ETS were tested and refined. Additionally, in the first two phases, the national governments were left in charge of planning the allocation of these certificates to their respective industries \citep{verde2019free}. These national allocation plans (NAP) were after the submittal of the national governments scrutinised by the European Commission \citep{ellerman2020}. Therefore, the verification procedure, including the submissions and verifications, of the NAPs, will be a significant source of the carbon policy surprise. Discussed further in Chapter \@ref(reggdates) and the exact dates, in addition to the shocks' nature, are specified in Appendix \@ref(regdates).

Phase II continued with a similar framework of submitted NAPs and their Commission approvals \citep{ellerman2020}. In the second phase, the governments were allowed to auction up to 10 per cent of the allowances. A raise from the 5 per cent in Phase I \citep{ellerman2020}. Industrial production slowed abruptly after the financial crisis, which made the cap non-binding, thus reducing the price of the allowances to near zero \citep{verde2019free}. The effects of the policy changes between the different phases on the prices of the emission trading allowances can be seen clearly in Figure \@ref(fig:price-evolution).

```{r price-evolution,echo=FALSE,message=FALSE,fig.cap="The evolution of the EU Emission trading system spot price and different future lengths prices through the different phases of the system. 2.0. Data was sourced from Datastream."}

library(tidyverse)
library(ggplot2)

get_eua_data()  %>%  
  select(1:5) %>% 
  pivot_longer(-date) %>%  
  ggplot(aes(x = date, y = value, colour = name)) +
  geom_line() +

  geom_vline(xintercept = c(lubridate::ymd('20071231'),
                            lubridate::ymd('20121231'),
                            lubridate::ymd('20201231')),
             linetype = "dashed", alpha = 0.5, size = 1)+
  labs(x = '', y = '€') +
  annotate(geom = "text",
           y = 80, 
           x = lubridate::ymd('20060630'), 
           label = 'Phase I')+
  annotate(geom = "text",
           y = 80, 
           x = lubridate::ymd('20100630'), 
           label = 'Phase II')+
  annotate(geom = "text",
           y = 80, 
           x = lubridate::ymd('20170630'), 
           label = 'Phase III')+
  annotate(geom = "text",
           y = 15, 
           x = lubridate::ymd('20211230'), 
           label = 'Phase IV')+
  scale_color_brewer(palette = 'Set1', 
                     name = '', 
                     labels = c('Settlement price',
                                'Front contract',
                                'Second contract',
                                'Third contract'))+
  theme_bw()+
  theme(legend.position = 'bottom')

```

The oversupply of ETS allowances in the latter stages of Phase II was the inspiration that led to the reforms in Phase III \citep{garcia2021}. The most substantial change to the ETS market was the abolition of NAPs and the resulting centralisation by adopting a single EU-wide cap \citep{verde2019free}. This cap was planned to reduce yearly by a linear amount that was decided to be 1.74 per cent of the 2010 total allowances. \citep{ellerman2020}. This linear decrease would lead to a 21 per cent reduction by 2020 in emissions in the sectors of the markets governed by the ETS compared to the levels in 2005 \citep{verde2019free}. Another major reform enacted in Phase III was phasing out the free allocation scheme to the energy sector in 2013, and plans to phase out the free allocation also in the remaining industrial sectors by 2027 \citep{ellerman2020}. The effects of these strict system overhauls can also be seen in figure \@ref(fig:price-evolution), where the news of future updates can be seen moving the futures price before it is realised at the spot price of the allowances. This is the essence behind the carbon policy surprise series and its usefulness in identifying the structural shocks in the SVAR model in chapter 5.

The changes that were enacted in Phase IV of the European Union ETS can be characterised by a more ambitious pace of allowance reductions and more stringent rules for the free allocations in the remaining sectors that still had them \citep{verde2019free}. The unintended consequences of this ambitious pace were countervailed by diverting the auction revenues to support energy sector modernisation by the innovation fund and modernisation fund \citep{verde2019free}. As Phase IV has only recently begun, the full implications of the rule changes are yet to be seen. Another challenge that will test the resolve of the European policy-makers is higher energy costs and the downstream effects of higher prices. I will bring more evidence in Chapter \@ref(results).

\newpage