```{r setupresult, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      fig.width = 7,
                      fig.height = 5,
                      fig.align = 'left')
library(ggplot2)
library(tidyverse)
library(readxl)
library(dplyr)

source('gradu_read_data.R', encoding = 'UTF-8')
```

# Results {#results}

In this chapter I will provide the the results of my research. In the first section I will provide the results of the first stage regression results of the power of the instrument. In the second section I will provide a short summary of the results from the research by @kaenzig2022. Third section is dedicated to the results of my thesis and fourth section is going to dwell into the reasons how these results may have arisen. 

## First stage {#firststage}

Inference with instrumental variables lies on the assumptions of relevancy with the shock of interest and exogeneity with other shocks. An other hidden assumption is that a weak correlation between the instrument and the structural shock of interest compromises the large sample validity of the inference based on the instrument. Due to this @montiel2021 propose that the first-stage heteroskedasticity-robust F-statistic between the instrument and the VAR-residual should be reported as a measure of the strength of the instrument. They also suggest a rule of thumb of $F > 10$ for to be sure that the instrument is not weak and  [@montiel2021].

For my model the $F = 10.99$, which would imply that the instrument is strong enough to be used with standard inference and there is no need to use the weak-instrument robust confidence sets that were elaborated further in @montiel2021. With the first stage regression we also find that the instrument explains $2.97\%$ of the energy price index residual, which is not . In conclusion there is no evidence to consider that a weak instrument problem would affect the following inference based on this instrument.

## Results of \citet{kaenzig2022} {#kanzigresult}

I will describe in this section the results of @kaenzig2022 as they differ in interesting ways to the results I bring forth in chapter \@ref(macroeff). This way the reader has a better understanding of the context of my results.

Firstly the instantaneous shock from the carbon policy shock to the energy component of the Harmonised Index of Consumer Prices (HICP) is normalised to the tune of one per cent. 6 months after the carbon policy shock the 90 per cent confidence interval is between $0.5$ and $2$ per cents. For the whole $50$ months, that the impulse response was estimated, the energy components of HICP stays statistically significantly above zero [@kaenzig2022].

The price increase in the energy prices effect the headline HICP. It has a instantaneous response of $0.1$ per cent and stays persistently, even though slightly declining, above zero [@kaenzig2022]. 

To GHG emissions of the EU area the carbon policy shock has a significant and persistent effect. Having a point estimate of $-0.5$ percent $12$ months after the carbon policy shock. Emission levels will slightly rise after the slump but are in a statistically significant manor below zero. Point estimate is roughly $-0.25$ after $50$ months [@kaenzig2022].

The response curve of the industrial output has a similar shape as does the GHG emissions. Initial slowdown that has a point estimate of around $-0.8$ and it occurs after $12$ months. In contrast with the GHG emissions the industrial output recovers to having a point estimate of zero after $40$ months. 

The unemployment rate responses to the carbon policy shock with a steady increase that peaks after $24$ months with a point estimate of $0.2$ percentage points. After the peak the unemployment rate stays steadily and statistically significantly above zero.

The response of the stock prices is negative in the short run, having a point estimate of $-2$ per cent, but with a wide confidence band that also covers zero. The stock prices rebound and after approximately $25$ months the point estimate is above zero [@kaenzig2022]. 

Policy rate of the three month Euribor rate has a substantial response to the carbon policy shock that is as its' deepest $0.1$ percentage points $20$ months after the carbon policy shock. It returns to zero after $40$ months [@kaenzig2022]. The shape of the impulse response of the policy rate seems to follow closely with the shape of the industrial policy. This is something that can also be seen in the figure \@ref(fig:irf).

Broad Real Effective exchange rate (REER) doesn't have a short run response to the carbon policy shock but after $15$ months there starts a statistically significant decline that persists till the end of the estimation period when the REER is $-0.6$ per cent [@kaenzig2022]. 

For more inquisitive examination of the results from the research done by @kaenzig2022 please refer to the Figure 3 of his paper [@kaenzig2022]. 

## Effects to Finnish macroeconomy {#macroeff}

In the following section I will present the resulting impulse response functions to a carbon policy shock. The solid black line is a point estimate and the darker and lighter regions are the 68 and 90 per cent confidence bands respectfully. These confidence bands are calculated with a moving block bootstrap which was first brought forth by @jentsch2019.^[The major advantage of the moving block bootstrap method, comparing it to the wild bootstrap method, is that it will produce accurate confidence intervals. As the wild bootstrap will produce inaccurate impulse response functions to SVAR-IV [@jentsch2016]. The code used in my thesis is relying on @kaenzig2021 reproduction files, which can be found from [ github](https://github.com/dkaenzig/replicationOilSupplyNews). I also inquired through email exchange that the methods used in that reproduction file are similar as in @kaenzig2022. If there are any mistakes to be found, they are naturally mine.] With a block size of 20 and with $10 000$ bootstrap replications.  

```{r irf, message=FALSE, warning=FALSE,  fig.width = 6 , fig.height=7.5, fig.cap='Impulse response functions from carbon policy shocks to macroeconomic indicators of Finland. The policy shock has been normalised to size of instantaneous 1 per cent rise in energy prices. Continuous line represents the point estimate, dark area  represents the 68 percent confidence intervals and the light are 90 percent. Variables 3 month euribor and unemployment rate are represented as percentage points, all others are percentage changes.'}

mat<-R.matlab::readMat('data/IRFsbench.mat')

logful <- get_log_data()

y <- logful[,c(5,15,4,16,3,17,7,14)]
mat$colnames <- colnames(y)

base<-cbind(c(0:50), mat$IRFs.base)
colnames(base) <- c('month', mat$colnames)
base <- tibble::as_tibble(base) %>%
  pivot_longer(-month) %>%
  mutate(type = 'base')

upper<-cbind(c(0:50), mat$IRFsupper.base)
colnames(upper) <- c('month', mat$colnames)
upper <- tibble::as_tibble(upper) %>%
  pivot_longer(-month) %>%
  mutate(type = 'upper')

lower<-cbind(c(0:50), mat$IRFslower.base)
colnames(lower) <- c('month', mat$colnames)
lower <- tibble::as_tibble(lower) %>%
  pivot_longer(-month) %>%
  mutate(type = 'lower')

upper2<-cbind(c(0:50), mat$IRFsupper2.base)
colnames(upper2) <- c('month', mat$colnames)
upper2 <- tibble::as_tibble(upper2)%>%
  pivot_longer(-month) %>%
  mutate(type = 'upper2')

lower2<-cbind(c(0:50), mat$IRFslower2.base)
colnames(lower2) <- c('month', mat$colnames)
lower2 <- tibble::as_tibble(lower2) %>%
  pivot_longer(-month) %>%
  mutate(type = 'lower2')

data<-base %>% 
  rbind(upper) %>% 
  rbind(lower) %>%
  rbind(upper2)%>%
  rbind(lower2)%>%
  pivot_wider(names_from = type)

data %>%
  mutate(across(name, factor, levels=mat$colnames)) %>%
  ggplot(aes(x = month))+
  geom_ribbon(aes(ymin = lower, ymax = upper),fill = '#3e99d6', alpha = 0.4)+
  geom_ribbon(aes(ymin = lower2, ymax = upper2),fill = '#3e99d6', alpha = 0.75)+
  geom_line(aes(y = base), size = 1) +
  geom_hline(yintercept = 0, linetype="dashed", alpha = 0.5)+
  theme_bw()+
  facet_wrap(~name, 
             ncol = 2, 
             scales = 'free_y',
             labeller = as_labeller(
               c(HCIP_energy = "Energy component of HICP",
               GHG_Känzig_method = "GHG emissions", 
               HCIP = "HICP",
               ind_adj = "Industrial output",
               Euribor_3kk = "3 month Euribor rate",
               unemp_adj = "Unemployment rate",
               omx_hel_25 = "OMX Helsinki 25 index",
               RBFIBIS = "Real Broad Effective Exchange Rate")
             ))+
  labs(y = '', x = 'Months')

```



As can be seen in the figure \@ref(fig:irf) the negative carbon policy shock is normalised to have a effect of $1$ per cent increase in the energy components of HICP. 

As the energy components, greenhouse gas emission, headline HICP, industrial output index,  OMX Helsinki 25 stock index and the real broad exchange rate index are all handled in log-levels can the plots be interpreted as percentual changes. In contrast the unemployment rate and the 3 month Euribor interest rate are handled in percentage points and thus the plots represent changes of percentage points.

What can be clearly seen in this figure is the high persistency of the higher prices as a response to the carbon policy shock. This persistency in higher energy prices is feeding into higher headline consumer price index and it is statistically significant until the end of the estimation period, $50$ months ahead. The behaviour of the consumer price indexes are inline with the findings of @kaenzig2022 findings from European wide context. Where my findings differ, and seem to defy the conventional wisdom, is the behaviour of the Finnish economy in the short-run. 

There is an rather unexpected, but statistically insignificant, evidence that the Finnish economy benefits from the carbon policy shocks in the short-run. This can be seen as the industrial output index which is slightly elevated after the shock. But then continues to decrease to a statistically significant long term negative impulse. Similar, but opposite signed, impulse is with the unemployment rate. Starting with a statistically insignificant negative simultaneous impact that peaks around 10 months and then steady rise to a statistically significant positive reaction. The point estimate would imply that a carbon policy shock that would instantaneously raise the energy prices by 1% would in $40$ months time lead to a $0.07$ percentage point rise in Finnish unemployment. 

There is also statistically significant evidence that the carbon policy shock affects the $3$ months Euribor interest rate, by reducing it with a point estimate of 0.1 percentage point in the long run. The effects of the carbon policy shock to the stock market seem to be statistically inconclusive. There is a noticeable evidence of effects to the Finnish real effective exchange rate in the long run.

## Discussion of the findings

The reasons behind these impulses are outside the bounds of this thesis. But for the sake of future research I will provide few hypotheses. I would suggest, that there are two different reactions at play. A short run reaction that stimulates the economy and a long run reaction that is slowing down the economy. One hypothetical reason behind the short run stimulation might arise from the long history of Finnish carbon taxes which as @elbaum2021 states made the Finnish industry more energy efficient. This energy efficiency would give Finnish industry a competetive advantage to its peers. Another possible explanation is the composition of Finnish industry^[According to [Statistics Finland](https://www.stat.fi/til/tti/2020/tti_2020_2021-07-01_tie_001_en.html) 45% of the value produced by Finnish industry is from metal industry products.]  with the high share of capital goods in the production basket. The short term stimulus emerge due to the other European countries investing to more sustainable capital which would, in turn make the products of Finnish industrial companies more attractive. These explanations are hardly exclusive and thus their relative share of effect or the existence of additional explanations will be left to future research.

The long term effects of these shocks are more in line with the findings of @kaenzig2022. This could be hypothesised to mean that the short term stimulus of some kind is running out and Finland will revert to the European norm. This would imply that the lower aggregate demand in Europe, which is demonstrated by \citeauthor{kaenzig2022}, would after the end of investment surge bring the Finnish economy to a lower long term level. Also as with the short term effects the long term will have to be left to the future research. 

\newpage