```{r setupresult, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      fig.width = 7,
                      fig.height = 5,
                      fig.align = 'left')
library(ggplot2)
library(tidyverse)
library(readxl)
library(dplyr)

source('gradu_read_data.R', encoding = 'UTF-8')
```

# Results {#results}

In this chapter, I will provide the results of my research. In the first section, I will provide the first-stage regression results of the instrument's power. In the second section, I will summarise the results from the research by @kaenzig2022 for the reader to better understand the context of the result of this thesis. The third section is dedicated to the results of my thesis, and the fourth section will dwell on why these results may have arisen. 

## First stage {#firststage}

Inference with instrumental variables lies in the assumptions of relevancy with the shock of interest and exogeneity with other shocks. However, another often unstated assumption is that a weak correlation between the instrument and the structural shock of interest compromises the large sample validity of the inference based on the instrument [@montiel2021]. Due to this, @montiel2021 proposes that the first-stage heteroskedasticity-robust F-statistic between the instrument and the VAR-residual should be reported as a measure of the instrument's strength. They additionally propose a rule of thumb of $F > 10$ to ensure the instrument is not weak and could be used with standard inference [@montiel2021]. If the instrument would be weak their research brings forth a method to use with weaker instrument variables.

For my model, the $F = 10.99$ which implies that the instrument is strong enough to be used with standard inference, and there is no need to use the weak-instrument robust confidence sets elaborated further in @montiel2021. With the first stage regression, we also find that the instrument explains $2.97\%$ of the energy price index residual, which is not. In conclusion, there is no evidence to consider that a weak instrument problem would affect the following inference based on this instrument.

## Results of \citet{kaenzig2022} {#kanzigresult}

The instantaneous shock from the carbon policy shock to the energy component of the Harmonised Index of Consumer Prices (HICP) is normalised to the size of one per cent. 6 months after the carbon policy shock the 90 per cent confidence interval is between $0.5$ and $2$ per cents. For the whole estimation period of $50$ months, the impulse response was estimated, the energy components of HICP stays statistically significantly above zero [@kaenzig2022].

The carbon policy shock would have effect also to the headline HICP. It has a instantaneous response of $0.1$ per cent and stays persistently, even though slightly declining, above zero [@kaenzig2022]. 

To GHG emissions of the EU area the carbon policy shock has a significant and persistent effect. Having a point estimate of $-0.5$ percent $12$ months after the carbon policy shock. Emission levels will slightly rise after the slump but are in a statistically significant manor below zero. Point estimate is roughly $-0.25$ after $50$ months [@kaenzig2022].

The impulse response function of the industrial output has a similar shape as does the GHG emissions. Initial slowdown that has a point estimate of around $-0.8$ and it also occurs after $12$ months. In contrast with the GHG emissions the industrial output recovers to having a point estimate of zero after $40$ months. 

The unemployment rate responses to the carbon policy shock with a steady increase that peaks after $24$ months with a point estimate of $0.2$ percentage points. After reaching the peak the unemployment rate stays steadily and statistically significantly above zero.

The impulse response function of the stock price index is negative in the short run, having a point estimate of $-2$ per cent, but with a wide confidence band that also covers zero. The stock prices rebound and after approximately $25$ months the point estimate is above zero [@kaenzig2022]. 

Policy rate of the three month Euribor has a substantial response to the carbon policy shock that is as its' deepest $0.1$ percentage points $20$ months after the carbon policy shock. It returns to zero after $40$ months [@kaenzig2022]. The shape of the impulse response of the policy rate seems to follow closely with the shape of the industrial policy. This is something that can also be seen in Figure \@ref(fig:irf), where the results of Finnish macroeconomy are reported.

Broad Real Effective exchange rate (REER) doesn't have a short run response to the carbon policy shock but after $15$ months there starts a statistically significant decline that persists till the end of the estimation period when the REER is $-0.6$ per cent [@kaenzig2022].

## Effects to Finnish macroeconomy {#macroeff}

In the following section, I will present the resulting impulse response to a carbon policy shock. The solid black line represents a point estimate, and the darker and lighter regions are the 68 and 90 per cent confidence bands, respectively. These confidence bands are calculated with a moving block bootstrap which was first brought forth by @jentsch2019.^[The significant advantage of the moving block bootstrap method, comparing it to the wild bootstrap method, is that it will produce accurate confidence intervals as the wild bootstrap will produce inaccurate impulse response functions to SVAR-IV [@jentsch2016]. The code used in my thesis relies on @kaenzig2021 reproduction files, which can be found from [ github](https://github.com/dkaenzig/replicationOilSupplyNews). I also inquired through email exchange that the methods used in that reproduction file are similar to @kaenzig2022. If any mistakes are found, they are naturally mine.] With a block size of 20 and with $10 000$ bootstrap replications.  

```{r irf, message=FALSE, warning=FALSE,  fig.width = 6 , fig.height=7.5, fig.cap='Impulse response functions from carbon policy shocks to macroeconomic indicators of Finland. The policy shock has been normalised to size of instantaneous 1 per cent rise in energy prices. Continuous line represents the point estimate, dark area  represents the 68 percent confidence intervals and the light are 90 percent. Variables 3 month euribor and unemployment rate are represented as percentage points, all others are percentage changes.'}

mat<-R.matlab::readMat('data/IRFsbench.mat')

logful <- get_log_data() 

y <- logful[,c(6,16,5,17,4,18,8,15)]
mat$colnames <- colnames(y)

base<-cbind(c(0:50), mat$IRFs.base)
colnames(base) <- c('month', mat$colnames)
base <- tibble::as_tibble(base) %>%
  pivot_longer(-month) %>%
  mutate(type = 'base')

upper<-cbind(c(0:50), mat$IRFsupper.base)
colnames(upper) <- c('month', mat$colnames)
upper <- tibble::as_tibble(upper) %>%
  pivot_longer(-month) %>%
  mutate(type = 'upper')

lower<-cbind(c(0:50), mat$IRFslower.base)
colnames(lower) <- c('month', mat$colnames)
lower <- tibble::as_tibble(lower) %>%
  pivot_longer(-month) %>%
  mutate(type = 'lower')

upper2<-cbind(c(0:50), mat$IRFsupper2.base)
colnames(upper2) <- c('month', mat$colnames)
upper2 <- tibble::as_tibble(upper2)%>%
  pivot_longer(-month) %>%
  mutate(type = 'upper2')

lower2<-cbind(c(0:50), mat$IRFslower2.base)
colnames(lower2) <- c('month', mat$colnames)
lower2 <- tibble::as_tibble(lower2) %>%
  pivot_longer(-month) %>%
  mutate(type = 'lower2')

data<-base %>% 
  rbind(upper) %>% 
  rbind(lower) %>%
  rbind(upper2)%>%
  rbind(lower2)%>%
  pivot_wider(names_from = type)

data %>%
  mutate(across(name, factor, levels=mat$colnames)) %>%
  ggplot(aes(x = month))+
  geom_ribbon(aes(ymin = lower, ymax = upper),fill = '#3e99d6', alpha = 0.4)+
  geom_ribbon(aes(ymin = lower2, ymax = upper2),fill = '#3e99d6', alpha = 0.75)+
  geom_line(aes(y = base), size = 1) +
  geom_hline(yintercept = 0, linetype="dashed", alpha = 0.5)+
  theme_bw()+
  facet_wrap(~name, 
             ncol = 2, 
             scales = 'free_y',
             labeller = as_labeller(
               c(HCIP_energy = "Energy component of HICP",
               GHG_Känzig_method = "GHG emissions", 
               HCIP = "HCIP",
               ind_adj = "Industrial output",
               Euribor_3kk = "3 month Euribor rate",
               unemp_adj = "Unemployment rate",
               omx_hel_25 = "OMX Helsinki 25 index",
               RBFIBIS = "Real Broad Effective Exchange Rate")
             ))+
  labs(y = '', x = 'Months')

```

As shown in Figure \@ref(fig:irf), the negative carbon policy shock is normalised to have a $1$ per cent increase in the energy components of HICP. 

The energy components, greenhouse gas emission, headline HICP, industrial output index,  OMX Helsinki 25 stock index and the real broad exchange rate index are all handled in log levels. The plots can therefore be interpreted as percentual changes. Only the unemployment and the 3-month Euribor interest rates are in percentage points.

What can be seen in this figure is the high persistence of the higher prices as a response to the carbon policy shock. This persistency in higher energy prices is feeding into the higher headline consumer price index, and it is statistically significant until the end of the estimation period, $50$ months ahead. The behaviour of the consumer price indexes is in line with the findings of @kaenzig2022 findings from a European-wide context. Where my findings differ from the results of @kaenzig2022, is the behaviour of the Finnish economy in the short-run. 

There is somewhat unexpected though statistically insignificant, evidence that some parts of the Finnish economy benefit from the carbon policy shocks in the short run. This can be seen as the industrial output index, which is slightly elevated after the shock. But then continues to decrease to a statistically significant long-term negative impulse response. A similar reaction is to the impulse of the unemployment rate. Starting with a statistically insignificant negative simultaneous impact that peaks around ten months and then a steady rise to a statistically significant positive reaction. The point estimate would imply that a carbon policy shock that would instantaneously raise the energy prices by 1% would, in $40$ months lead to a $0.07$ percentage point rise in Finnish unemployment. 

There is also statistically significant evidence that the carbon policy shock affects the $3$ month Euribor interest rate by reducing it with a point estimate of 0.1 percentage point in the long run. The effects of the carbon policy shock on the stock market are statistically inconclusive. There is noticeable evidence of effects on the Finnish real effective exchange rate in the long run.

## Discussion of the findings

The causal reasoning behind these impulses is not the topic of this thesis. Nevertheless, for the sake of future research, I will entertain a few hypotheses that might provide explanatory value. First, I would suggest that two different reactions are at play. A short-run reaction stimulating the economy, and a long-run reaction slowing down the Finnish economy. One hypothetical reason behind the short-run stimulation is the long history of Finnish carbon taxes, which as @ elbaum2021 states, made the Finnish industry more energy efficient. This energy efficiency would give the Finnish industry a competitive advantage over its peers. Another possible explanation is the composition of Finnish industry^[According to [Statistics Finland](https://www.stat.fi/til/tti/2020/tti_2020_2021-07-01_tie_001_en.html) 45% of the value produced by Finnish industry is from metal industry products.]  with the high share of capital goods in the production basket. The short-term stimulus emerged due to the other European countries investing in more sustainable production capital, which would make Finnish industrial companies' products more attractive. These explanations are hardly exclusive, and thus their relative share of effect or the existence of additional explanations will be left to future research.

The long-term effects of these shocks are more in line with the findings of @kaenzig2022. This could be hypothesised to mean that the short-term stimulus of some kind is running out, and Finland will revert to the European norm. This, in turn, would imply that the lower aggregate demand in Europe, demonstrated by \citeauthor{kaenzig2022}, would bring the Finnish economy to a lower long-term level after the investment surge. Also, as with the short-term effects, the long-term will have to be left to future research. 

\newpage