

```{r setuptable, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      fig.width = 7,
                      fig.height = 5,
                      fig.align = 'left')
library(ggplot2)
library(tidyverse)
library(readxl)
library(dplyr)

source('gradu_read_data.R', encoding = 'UTF-8')
```


# (APPENDIX) Appendix {-} 

# Data sources {#datasaources}

The data has been collected from various 

\newpage

# Regulatory dates {#regdates}

In this appendix I will detail the dates from previous research and the dates that I have identified using the same sources and methodology as @kaenzig2022.  

The first stage of regulatory news which happened prior to year $2010$ were retrieved by \citeauthor{kaenzig2022} the from the [official journal of the European Union](https://eur-lex.europa.eu/homepage.html) [@kaenzig2022]. \citeauthor{kaenzig2022} also added the information of the dates of the NAP decisions which was collected from the research of @mansanet2009impacts which covers the phase I and phase II [@kaenzig2022, @mansanet2009impacts]. After the year $2010$ dates of interest are archived to the [European Commission Climate Action news archive](https://ec.europa.eu/clima/news/news_archives_en) [@kaenzig2022]. All of these dates are collected to the table \@ref(tab:kanzig-reg). 

I followed the example of @kaenzig2022 and collected the regulatory dates also from the [European Commission Climate Action news archive](https://ec.europa.eu/clima/news/news_archives_en) and these dates are detailed in the table \@ref(tab:my-reg).

\begingroup
\singlespacing

```{r kanzig-reg, warning=FALSE}

readxl::read_xlsx('data/kaenzig_reg_dates.xlsx') %>% 
  kableExtra::kbl(caption = 'Regulatory dates from the research of \\citet{kaenzig2022}', 
               booktabs = TRUE,
               longtable = TRUE) %>%
  kableExtra::column_spec(2, width = "35em") %>% 
  kableExtra::kable_styling(position = "left",
                            latex_options = c("hold_position", "repeat_header", "scale_down"), 
                            full_width = F,
                            font_size = 10)

```

\newpage

```{r my-reg, warning=FALSE}

readxl::read_xlsx('data/my_reg_dates.xlsx') %>% 
  kableExtra::kbl(caption = 'Regulatory dates I have found using similar methodology as \\citet{kaenzig2022}.', 
               booktabs = TRUE,
               longtable = TRUE) %>%
  kableExtra::column_spec(2, width = "35em") %>% 
  kableExtra::kable_styling(position = "left",
                            latex_options = c("hold_position", "repeat_header", "scale_down"), 
                            full_width = F,
                            font_size = 10)

```


\endgroup
\newpage